---
title: "Homework 5.A"
author: 'Papadopoulos Christos'
output: flexdashboard::flex_dashboard
runtime: shiny
---

```{r global, include=FALSE}
# load data in 'global' chunk so it can be shared by all users of the dashboard
library(tidyverse)
library(cshapes)
library(sf)
library(tidyverse)
library(tidygraph)
library(terra)
library(spData)
library(spDataLarge)
library(tmap)
```
Page 1
===================================== 

Column {.sidebar}
-----------------------------------------------------------------------

Distance of countries with Greece in km

```{r}
selectInput("dist", label = "Distance(km):",
            choices= c(1000,1500,2000,2500,3000),selected = 2000)
```

Column
-----------------------------------------------------------------------

### Map of countries close to Greece

```{r,include=TRUE}
cmap.2019 <- cshp(date=as.Date("2019-1-1"),useGW = TRUE)
sf_use_s2(FALSE)

```


```{r}
renderPlot({
  dist <-distlist(as.Date("2019-1-1"),keep=0.1, useGW = TRUE, type= 'mindist')
  dist_greece <- dist %>%
    filter(ccode1== 350)%>%
    filter(mindist < as.numeric(input$dist))
  countries_close_greece <- cmap.2019%>%
    filter(gwcode %in% dist_greece$ccode2)
  plot(countries_close_greece$geometry,
     main= 'Countries close to Greece')
})
```

Page 2
===================================== 
Column {.sidebar}
-----------------------------------------------------------------------

```{r}
selectInput("year", label = "Year of interest:",
            choices= c('1900-1-1','1920-1-1','1940-1-1','1960-1-1','1980-1-1','2000-1-1','2019-1-1'),selected = '2019-1-1')
```
Column
------------------------------------------------------------------------
### Map of countries and their capitals

```{r}
renderPlot({
  date= toString(input$year)
 
  map= cshp(date=as.Date(date),useGW = TRUE)
  cap_loc= cbind(map$caplong,map$caplat)
  cap_loc_st = st_multipoint(cap_loc)
  cap_loc_sfc= st_sfc(cap_loc_st, crs= 'EPSG:4326')

  capitals <- map$capname
  capitals_data <- st_sf(data.frame(Capitals=capitals),geom = cap_loc_sfc)
  plot(map$geometry,
     main= 'Capitals of the world')
  plot(capitals_data,add= TRUE, col='red',cex=0.5)
  text(jitter(cap_loc),capitals)
})

```
Page 3
===================================== 
Column {.sidebar}
-----------------------------------------------------------------------

```{r}
selectInput("buf", label = "Buffer Length (km):",
            choices= c(50,100,150,200),selected = 100)
```
Column
------------------------------------------------------------------------
### Map of countries that the buffer of their capitals include another's country territory 

```{r}
renderPlot({
  sf_use_s2(TRUE)
  countries <- cmap.2019 %>%
    rowwise()%>%
    mutate(capitals_coordinates = st_sfc(st_point(c(caplong,caplat)),crs='EPSG:4326' ))%>%
    ungroup()
  buffer <- as.numeric(input$buf) *1000 
  capitals_buffer <- st_buffer(countries$capitals_coordinates,dist=buffer)
  capitals_overlapping <- st_overlaps(capitals_buffer, cmap.2019)
  capitals_overlapping_logical <- lengths(capitals_overlapping) >0
  plot(countries[capitals_overlapping_logical,]$geometry)
})
```
Page 4
=====================================

###  Distance of capitals from the centroids
Ranking of  the countries according to the distance of their capitals from the
centroids of their polygons.The map of the world where the countries 
are classified into three categories: 1 for distances <100, 2 for distances
[100,300] and 3 for distances >300.

```{r}
renderPlot({
  countries <- cmap.2019 %>%
    rowwise()%>%
    mutate(capitals_coordinates = st_sfc(st_point(c(caplong,caplat)),crs='EPSG:4326' ))%>%
    ungroup()
  countries_centroids <- countries%>%
    rowwise()%>%
    mutate(centroids = st_centroid(geometry, of_largest_polygon = TRUE))%>%
    ungroup()%>%
    rowwise()%>%
    mutate(dist_cap_cent= st_distance(centroids,capitals_coordinates))%>%
    ungroup()%>%
    arrange(-dist_cap_cent)
  library(units)
  countries_centroids$dist_cap_cent <- ifelse(countries_centroids$dist_cap_cent<set_units(100,km),'1',
                                            ifelse(countries_centroids$dist_cap_cent<=set_units(300,km),'2','3'))
  plot(countries_centroids %>% select(dist_cap_cent),main= 'Distance centroid from the capitals')

})
