---
title: "Homework 5.B"
author: "PAPADOPOULOS CHRISTOS"
date: "2023-01-15"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Import necessary libraries and create a proximity matrix between the coutries
```{r}
library(tidyverse)
library(cshapes)
library(sf)

cmap.2019 <- cshp(date=as.Date("2019-1-1"))
validcmap <- st_make_valid(cmap.2019)
capitals_distance <- distmatrix(date= as.Date('2019-1-1'),type='capdist')
capitals_distance<- as.matrix(capitals_distance)
# βιβλιοθήκη που δημιουργεί proximity matrices
library(prioritizr)

#create proximity matrix
prox_matrix <- proximity_matrix(validcmap,0)
prox_matrix <- as.matrix(prox_matrix)

```

```{r}
df= NULL
for (i in 1:nrow(validcmap)){
  from_capital = validcmap$capname[i]
  for (j in 1:ncol(prox_matrix)){
    if (prox_matrix[i,j]==1){
      to_capital= validcmap$capname[j]
      distance= capitals_distance[i,j]
      df= rbind(df, data.frame(from_capital,to_capital,distance))
    }
  }
}

head(df)
```

```{r}
library(tidygraph)
graph_df <- as_tbl_graph(df)
graph_df

```

```{r}
library(stringr)
graph_df <- graph_df%>%
  activate(nodes) %>%
  mutate(
    title = str_to_title(name),
    label = str_replace_all(title, " ", "\n")
  )
graph_df
```

```{r}
capitals <- graph_df%>%
  activate(nodes)%>%
  pull(title)

```

### Provide a graph with the network you created (experiment with the various ggraph layouts for the aesthetically best result for you). In in any case to display at least the names of the nodes. You can use the distance between two nodes to color or set the thickness of the edges that connect them.

```{r}
# Visualizing
library(ggraph)
graph_df %>%
  ggraph(layout = "kk") +
  geom_node_text(aes(label = label), size = 1.5,repel=TRUE) +
  geom_edge_diagonal(aes(color=distance), alpha = 0.4)

```


### Shortest path between 'Athens' and 'Copenhagen' when the weight is the distance between capitals


```{r}
from <- which(capitals=='Athens')
from
to <- which(capitals=='Copenhagen')
to
shortest <- graph_df %>%
  morph(to_shortest_path,from,to,weights= distance)

shortest <- shortest %>%
  mutate(selected_node = TRUE) %>%
  activate(edges) %>%
  mutate(selected_edge = TRUE) %>%
  unmorph() 
shortest <- shortest %>%
  activate(nodes) %>%
  mutate(selected_node = ifelse(is.na(selected_node), 1, 2)) %>%
  activate(edges) %>%
  mutate(selected_edge = ifelse(is.na(selected_edge), 1, 2)) %>%
  arrange(selected_edge)
shortest %>%
  ggraph(layout = "kk") +
  geom_edge_diagonal(aes(alpha = selected_edge), color = "gray") +
  geom_node_text(aes(label = label, alpha = selected_node ), size = 2)

```


### Shortest path between 'Athens' and 'Copenhagen' when the weight is the number of total nodes visited.

```{r}
from <- which(capitals=='Athens')
from
to <- which(capitals=='Copenhagen')
to
shortest <- graph_df %>%
  morph(to_shortest_path,from,to,weights= NULL)
shortest <- shortest %>%
  mutate(selected_node = TRUE) %>%
  activate(edges) %>%
  mutate(selected_edge = TRUE) %>%
  unmorph() 
shortest <- shortest %>%
  activate(nodes) %>%
  mutate(selected_node = ifelse(is.na(selected_node), 1, 2)) %>%
  activate(edges) %>%
  mutate(selected_edge = ifelse(is.na(selected_edge), 1, 2)) %>%
  arrange(selected_edge)
shortest %>%
  ggraph(layout = "kk") +
  geom_edge_diagonal(aes(alpha = selected_edge), color = "gray") +
  geom_node_text(aes(label = label,  alpha = selected_node ), size = 2) 

```