---
title: "Homework 2 ggplot2"
author: "Papadopoulos Christos"
date: "2022-11-02"
output: html_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1
```{r , include=FALSE}
library(tidyverse)
library(ggplot2)
path = "C:/Users/Chrispdl/Desktop/edav/hw2/mcdonalds.csv"
mac <- read.csv(path)
```

### Preview the data 

```{r}
head(queen) 
summary(queen)
```
 
###  1a) Bar plot num of tracks per album
```{r}
barplot<-ggplot(data=queen, aes(x=album_name)) +
  geom_bar()+
  labs(title='Bar plot of tracks per album')+
  theme(axis.text.x = element_text(angle = 45,hjust=1))
barplot
```

### 1b)  Histogram of tempo distribution
```{r}
histogram <- ggplot(data=queen, aes(x=tempo))+
  geom_histogram(bins=15,aes(fill=mode))+
  labs(title='Histogram of Tempo distribution')
histogram
``` 

### 1c) Boxplot of tempo per album

```{r}
boxplot <- ggplot(data=queen, aes(x=album_name, y= tempo))+
  geom_boxplot(fill='lightblue')+
  labs(title='Boxplot of Tempo per album')+
  theme(axis.text.x = element_text(angle = 45,hjust=1))
boxplot
```

### 1d) Scatterplots of danceability and energy per album

``` {r}
scatter <- ggplot(data=queen, aes(x=energy, y=danceability))+
  geom_point(color="lightblue",aes(size=tempo)) + 
  facet_wrap(~ album_name)+
  labs(title='Scatterplots of danceability and energy per album')
  
scatter  

```

### 1e) 2 Plots of my choice
```{r}
scatter2 <- ggplot(data=queen, aes(x=danceability, y=track_popularity))+
  geom_point(color="lightblue",aes(size=tempo)) + 
  labs(title='Scatterplot of danceability and track popularity')
scatter2

scatter3 <- ggplot(data=queen, aes(x=valence, y=energy))+ 
  geom_point(aes(color=track_popularity))+
  labs(title='Scatterplot of valence and energy')
  
scatter3
```

# Exercise 2
```{r include=FALSE}
mac<- read.csv(file = 'mcdonalds.csv')
```

###Preview the data 
```{r}
head(mac)
summary(mac)
```

### Create a new dataframe with only complete observations
```{r}
complete_mac <- mac[complete.cases(mac),]
```

### Create a new column with price of Big Mc in dollars $
```{r}
final_mac <- complete_mac %>%
  transmute(country=name, price= local_price/dollar_ex,GDP_dollar= GDP_dollar)
```


### Scatterplot of  countries' GDP and price of Big Mc in dollars
```{r}
country_GDP_price_of_bigMc <- ggplot(data= final_mac, aes(x=GDP_dollar ,y=price))+
  geom_point()+
  labs(title='Scatterplot GDP - price of BigMac')
country_GDP_price_of_bigMc 
```


