---
title: "Exploring 120 Years of Olympic History"
author: "Christos Papadopoulos"
subtitle: Exploratory Data Analysis using R
output:
  html_document:
    df_print: paged
---

## Introduction 

Exploratory data analysis (EDA) is not a formal process but rather the investigation of many ideas that occur to the analyst at the initial phase of working with a dataset. It helps to understand the dataset better. Some of the initial ideas will end up as dead ends, while some other ones might lead to interesting insights. Exploratory data analysis is an important part of data analysis, it is done through generating questions about the data, searching for answers by visualizing, transforming, and modeling the data, and then using what could be learned from the previous steps to refine the questions or creating new ones. The main goal of EDA is to create many interesting leads that can be explored more in-depth at the later phase of data analysis (R for Data Science).

There are many analytical techniques that allow us to convert data into information. Descriptive analysis tries to answer the question “what happened in the past?”, inferential statistics asks “what about the rest of the population?”, diagnostic analysis tried to uncover what are the reasons for certain trends, predictive analysis answers the question “what’s likely to happen next?”, and prescriptive analysis answers the question “what should we do about it?” (Jones, 2020).

This project will be focused on the analysis of a dataset regarding 120 years of the Olympics. The dataset contains 2351 observations(rows) and 8 variables (columns) with information about the Olympic Games  held between 1896 - 2016.The first modern Games was held in Athens in 1896.

## Preview the data

```{r }
library(tidyverse)
path <- "C:/Users/Chrispdl/Desktop/edav/hw4/results.csv"
olympics <- read.csv(path)
#preview the data
olympics
summary(olympics)

```
##Cleaning the data
```{r}
olympics<- na.omit(olympics)
b<-table(olympics['Gender'])
b

#replace h in result column with :
olympics$Result <- str_replace(olympics$Result,'h',':')
olympics$Result <- str_replace(olympics$Result,'P','')
olympics
#delete rows with none in result column (but first check how many and if they are random)
x <- olympics%>%
  filter(Result == 'None')%>%
  group_by(Year)%>%
  summarize(count= n())

y<- olympics%>%
  group_by(Year)%>%
  summarize(count= n())



olympics
olympics <-olympics %>%
  filter(Result != 'None')
olympics

#check number of distinct events
num_events <- olympics%>%
  group_by(Event)

num_events

#check if ':' is present in olympics$Result
#grepl(':',olympics$Result)

#separate the original data in 2 dfs 
#result in time
olym_1 <- olympics%>%
  filter(grepl(':',olympics$Result)| Event== '100M Men'| Event=='110M Hurdles Men' |Event=='200M Men'| Event== '400M Hurdles Men'|Event== '400M Hurdles Women'| Event=='400M Men' |Event=='4X100M Relay Men'| Event=='100M Hurdles Women' | Event=='100M Women'| Event=='200M Women' | Event=='400M Hurdles Women
'| Event== '400M Women'| Event=='4X100M Relay Women' )

#result not in time 
olym_2 <- olympics%>%
  filter(!(grepl(':',olympics$Result)|Event=='100M Men'| Event=='110M Hurdles Men' |Event=='200M Men'| Event== '400M Hurdles Men'|Event== '400M Hurdles Women'|Event=='400M Men' |Event=='4X100M Relay Men'| Event=='100M Hurdles Women' | Event=='100M Women'| Event=='200M Women' | Event=='400M Hurdles Women
'| Event== '400M Women'| Event=='4X100M Relay Women'))

#num of events in the original data 
events <- olympics%>%
  summarise(events= n_distinct(Event))
events
#events in the olym_1 df
unique(olym_1$Event)

#events in the olym_2 df
unique(olym_2$Event)
```

## MEN AND WOMEN MEDALISTS
In this part , I would like to see the gender ratio on the medalists of Olympic games.
```{r}
plot1 = barplot(table(olympics$Gender), 
            main = "How many Olympic medalists are male and female?", 
            xlab = "Gender",
            ylab = "Number of athletes",
            ylim = c(0,2000),
            col = c("lightcoral", "cornflowerblue"))
plot1
```
## Countries with the most medals
In this part, I would like to see which country obtained the most gold medals in each Olympic Games.
```{r}
b<-table(olympics['Gender'])
b
olympics
c<-table(olympics['Year'])
c
fig_dat7<-olympics %>% filter(Medal=='G') %>%
  group_by(Year,Nationality) %>%
  summarise(Gold_Medal_count=n_distinct(Event))%>%
  top_n(1, Gold_Medal_count) %>% arrange(desc(Gold_Medal_count))

fig_dat7

fig_dat7 %>% ggplot(aes(x=factor(Year), y=Gold_Medal_count, fill=Nationality)) + 
  geom_bar(stat='identity',position =   position_dodge()) + 
  geom_text(aes(x=factor(Year), y=Gold_Medal_count/2,label=Nationality),size=2, position =position_dodge(width=1)) +
  geom_text(aes(x=factor(Year), y=Gold_Medal_count*1.05,label=Gold_Medal_count),size=2, position =position_dodge(width=1)) +
  coord_flip() +
  labs(x="Year",y="Gold Medals")

#Exploring the number of medals won by each country

winners <- olympics%>%
  group_by(Nationality,Medal, Year)%>%
  summarise(count= length(Medal))

#ordering countries by the total medal count and picking top 10
top_10 <- winners %>%
  group_by(Nationality)%>%
  summarize(Total_medals = sum(count))%>%
  arrange(desc(Total_medals))%>%
  slice(1:10)


#plot top_10
plot_3 <- top_10 %>%
  ggplot(aes(reorder(Nationality, -Total_medals),Total_medals))+
  geom_col()+
  labs(x= "Country code", y = "Total number of medals", title="Top 10 countries by the total number of medals") +
  theme_minimal()+
  scale_x_discrete(guide = guide_axis(angle = 90)) 
plot_3  
```

We see that USA is the country with the most medals most of the years.Another interesting thing is that we have only few observation for the olympic games of 1988.

## GREECE
Let's see what colors of the medals the team GREECE  scored over the years
```{r}
Greece1 = dplyr::filter(olympics, Nationality == "GRE")%>%
  group_by(Medal)%>%
  summarize(Count= n())

plot_4 <-ggplot(Greece1, aes(x = Medal, y = Count, fill = Medal)) +
  geom_bar(stat="identity") + 
  theme_minimal() + 
  labs(x = " ", y = "Total number of medals", title="Number of medals of each color for Greece") +
  scale_fill_manual(values=c("saddlebrown", "gold", "gray62"))
plot_4
```
Greece men vs women
```{r}
Greece2 = dplyr::filter(olympics, Nationality == "GRE")%>%
  group_by(Gender)%>%
  summarize(Count=n())
plot_5 <-ggplot(Greece2, aes(x = Gender, y = Count, fill = Gender)) +
  geom_bar(stat="identity") + 
  theme_minimal() + 
  labs(x = " ", y = "Total number of medals", title="Number of medals of each color for Greece") +
  scale_fill_manual(values=c("lightblue", "pink"))
plot_5
```

## 100 MEN
I would like to indentify how much the time has been reduced and which countries' athletes are the best 
Lets create a plot of 100M Golden Medalists Throught the years

```{r}
men_100 <- olym_1 %>%
  filter(Medal == 'G', Event=='100M Men')
men_100$Result <- as.numeric(men_100$Result)


best_time <- min(men_100$Result)
plot_6 <- ggplot(men_100, aes(x=Year,y=as.numeric(Result), color=Nationality))+
  geom_point(size=4)+
  scale_y_continuous(breaks=seq(9,12,0.1)) +
  labs(x = "Year of Olympic Games", y = "Seconds") +
  ggtitle("ScatterPlot of 100M Men Golden Medalists Throughout The Years")
plot_6
```

## The THORS of olympic games 
Scatter plot for hammer throw throughout the years
```{r}
hammer_men <- olym_2 %>%
  filter(Event=='Hammer Throw Men')%>%
  select(Gender,Year,Result)
max_ham_men<- max(hammer_men$Result)
min_ham_men <- min(hammer_men$Result)



hammer_women<-olym_2%>%
  filter(Event=='Hammer Throw Women',Gender=='W')%>%
  select(Gender,Year,Result)

max_ham_women<- max(hammer_women$Result)
min_ham_women <- min(hammer_women$Result)


hammer <- rbind(hammer_men,hammer_women)


#plot9
plot9 <- ggplot(hammer,aes(x=Year,y=as.numeric(Result),color=Gender))+
  geom_point()+
  scale_y_continuous(breaks=seq(35,100,5))+
  scale_colour_manual(values=c("blue", "red")) +
  ggtitle("ScatterPlot of Men/Women Hammer Throw Throughout The Years")+
  theme(legend.position = c(0.91, 0.08))+
  labs(x = "Year of Olympic Games", y = "Throw (in meters)") +
  geom_text(aes(label = ifelse((Result) %in% 
                                 c(max_ham_men, max_ham_women),
                               paste("Highsest: ", as.character(Result), "→", sep = " "),'')),
            hjust = 1.0, vjust = 0.5, col = "grey35") +
  geom_text(aes(label = ifelse((Result) %in% 
                                 c(min_ham_men,min_ham_women),
                               paste("←", as.character(Result), ":Lowest", sep = " "),'')),
            hjust = -0.12, vjust = 0.35, col = "grey35")
plot9
```

We see that Women's hammer throw was first included in the Olympics at the 2000 .Men's and women's best results are very close at this sport.

## Olympic Records
```{r}
records_1<- olym_1%>%
  group_by(Event)%>%
  summarize(min=min(Result),country=Nationality[which(Result==min(Result))])

olym_2$Result<-as.numeric(olym_2$Result)
records_2 <-olym_2%>%
  group_by(Event)%>%
  summarize(max=max(Result),country=Nationality[which(Result==max(Result))])

olympic_records<- rbind(records_1,records_2)%>%
  group_by(country)%>%
  summarize(count=n())

plot10 <- olympic_records%>%
  ggplot(aes(reorder(country,-count),count))+
  geom_col()+
  scale_y_continuous(breaks=seq(0,15,1))+
  labs(x= "Country code", y = "Total number of olympic records ", title="Countries which hold olympic record",subtitle='Best of the best')+
  scale_x_discrete(guide = guide_axis(angle = 90))
plot10
```
