---
title: "Homework 3"
author: "Papadopoulos Christos"
date: "2022-10-27"
output: html_document
---
## 1A)
#### Download and preview the data
```{r }
library(tidyverse)
dict_url <- "https://raw.githubusercontent.com/edwindj/datacleaning/master/data/dirty_iris.csv"
if (!file.exists('dirty_iris.csv'))download.file(dict_url,'dirty_iris.csv')
data <- read_csv('dirty_iris.csv')
```

#### Check that strings are not converted to factor.
```{r}
data
str(data)
is.factor(data)

```

## 1B)

#### Calculate the number and percentage of observations that are complete
```{r}
data_na=apply(X = is.na(data), MARGIN = 1, FUN = sum)
data_na_sum <- sum(data_na != 0)  
data_na_sum
nrow(data)
perc_compl_obser= (nrow(data)-data_na_sum)/nrow(data)
perc_compl_obser

```

## 1C)

#### Firstly i create a function is.special that finds special values and then i replace them with NA

```{r}
is.special <- function(x){
  if (is.numeric(x)) !is.finite(x) else is.na(x)
}

data[sapply(data, is.special)== TRUE] <- NA
```

## 2A)

#### Read the rules with editfiles 
```{r include=FALSE}
library(deducorrect)
library(editrules)
```
```{r}
E <- editfile("rules1.txt")
E
```
## 2B)
#### With violatedEdits i summarize how often each rule is broken 
```{r}
ve <- violatedEdits(E, data) #violated edits
summary(ve)
```


## 2C)
#### We filter the result of violatedEdits to find how often no rule is broken
```{r}
errors <- summary(ve)
zeros_errors <- errors %>%
  filter(errors==0)%>%
  summarise(num_of_good_observations=freq)
zeros_errors
```

## 2D)
#### We do the same to find out which observations have too long petals
```{r}
df_ve <- as.data.frame.table(ve)
colnames(df_ve) <- c('record','constraint','type')
long_petals<- df_ve %>%
  filter(constraint == 'num6',type== TRUE)%>%
  summarise(record)
long_petals
```

## 2E)
#### By using boxplot we find outliers in sepal length and then we set the outliers to mean of the Sepal.Length column.

```{r}
boxplot.stats(data['Sepal.Length'])$out
boxplot(data['Sepal.Length'],ylab='Sepal Length',notch=TRUE, varwidth= TRUE, main='Sepal length Boxplot')

data$Sepal.Length[(data$Sepal.Length==73) |(data$Sepal.Length==0)|(data$Sepal.Length==49)] <-mean(data$Petal.Length,na.rm=TRUE)
boxplot.stats(data['Sepal.Length'])$out

```

## 3A)
#### Replace non positive values from Petal.Width with NA using correctWithRules 
```{r}
rules <- correctionRules(expression(if (is.na(Petal.Width)) Petal.Width<- NA 
                                    else if(Petal.Width < 0) Petal.Width <-NA))
x<-correctWithRules(rules,data)
data <- x$corrected

```

## 3B)
#### With localizeErrors we create a logical Dataframe which shows every observation and if it breaks any rule and then we replace all the true values in the original dataframe with NA.
```{r}
le <-  localizeErrors(E, data)$adapt #localize errors
le
data[le==TRUE] <- NA

```

## 4A)
#### Use kNN imputation (VIM) to impute all missing values
```{r}
library(VIM)
data_KNN <- kNN(data)%>%
  select(1:5)
```


## 4B)
#### Use sequential hotdeck imputation to impute Petal.Width by sorting the dataset on Species.
```{r}
data_hot1 <- hotdeck(data=data, variable= 'Petal.Width', ord_var= 'Species')%>%
  select(1:5)
library(diffdf)
```

#### Compare the imputed Petal.Width with the sequential hotdeck imputation method.

```{r}
compare_hot1_knn <-diffdf(data_hot1, data_KNN)$VarDiff_Petal.Width
colnames(compare_hot1_knn) <- c('Variable','Diff_in_row','hotdeck_imput','knn_imput')
compare_hot1_knn<- compare_hot1_knn%>%
  mutate(abs_diff= abs(knn_imput - hotdeck_imput))
compare_hot1_knn
```

## 4C)
#### We do the same as before but now we sort the dataset on Species and Sepal.Length.
```{r}
data_hot2 <- hotdeck(data=data, variable= 'Petal.Width', ord_var= c('Species','Sepal.Length'))%>%
  select(1:5)
compare_hot2_knn <-diffdf(data_hot2, data_KNN)$VarDiff_Petal.Width
colnames(compare_hot2_knn) <- c('Variable','Diff_in_row','hotdeck_imput','knn_imput')
compare_hot2_knn<- compare_hot2_knn%>%
  mutate(abs_diff= abs(knn_imput - hotdeck_imput))
compare_hot2_knn
```


